### 2022-06-09 H - Posterior Distributions Tables/Online Supplement H - Tables with Information on Posterior Distributions.Rmd

---
title             : "Online Supplement H: Tables with Information on Posterior Distributions"
author            : " "
bibliography      : ["r-references.bib"]

header-includes   :
   - \usepackage{bm}
   - \usepackage{amsmath}
   - \usepackage{setspace}
   - \usepackage{xcolor}
   - \definecolor{mypink}{RGB}{255, 230, 255}
   - \usepackage{todonotes}
   - \usepackage{graphicx}
   - \usepackage{float}
   - \renewcommand{\tablename}{Table G}
   - \makeatletter
   - \def\tnum@table{\tablename\thetable}
   - \makeatother

csl               : apa6.csl
output            : pdf_document
---
\spacing{1.5}
\fontsize{12}{12}

```{r setup, include=FALSE}
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})

knitr::opts_chunk$set(tidy.opts=list(width.cutoff=40), tidy=TRUE, echo = TRUE, message = FALSE, warning = FALSE, fig.pos = "H")
```

\scriptsize 
```{R loadpackages, message = FALSE, warning = FALSE, echo = FALSE}
# Load the packages
library("papaja")
library("LaplacesDemon")
library("rstan")
library("brms")
library("plyr")
library("lattice")
library("ggplot2")
library("dplyr")
library("readr")
library("rmarkdown")
library("Rmisc")
library("devtools")
library("gghalves") 
library("bayesplot")
library("bridgesampling") 
library("gridExtra") 
library("tibble") 
library("kableExtra") 
library("bayestestR")
library("ggridges")
library("truncnorm")
library("ggbeeswarm")
```
\normalsize

|    This Online Supplement contains information on the posterior distributions of the general effects and individual variation estimated by `rstan` and `brms` for the normal and log-normal models. In the first table, the estimated mean of the posterior distribution, the standard error of this mean, the lower and upper bound of the 95% credible interval, the $\hat{R}$ and the number of effective samples are provided. In the second table, the estimated variance of the posterior distribution, the standard error of this variance, the lower and upper bound of the 95% credible interval, the $\hat{R}$ and the number of effective samples are provided. 
&nbsp;

## Normal model 
### Rstan 
\normalsize 
```{r loadmodelfit1, echo = FALSE}
setwd('..')
model_fit <- readRDS(file = "R objects/rstan/normal model/hier_modelc_adj_07012022.rds")
```

```{r geneftable1, results = "asis", cache = TRUE, echo = FALSE}
# Obtain the mean, standard error, 95% credible interval, rhat and number of effective samples for the intercept, and the digit and side parameters 
fitrstanhiertable <- summary(model_fit,  # model fit 
                             # Select parameters 
                             pars = c("mu", "mu2", "mu3", "mu4", "mu5", 
                                      "mu6"))$summary[,c("mean", "se_mean", 
                                                         "2.5%", "97.5%", 
                                                         "n_eff", "Rhat")]

# The results are transformed to a dataframe that allows us to plot the results nicely in a table 
fitrstanhiertable2 <- as.data.frame(fitrstanhiertable)

# We manually set the row names of the table using latex for mathematical symbols 
rownamesfitrstanhiertable <-  c("$\\mu_{\\gamma}$","$\\mu_{\\beta}$", "$\\mu_{\\delta_{7}}$", "$\\mu_{\\delta_{6}}$", "$\\mu_{\\delta_{4}}$", "$\\mu_{\\delta_{3}}$")

# We manually set the column names of the table 
colnames(fitrstanhiertable2) <- c("Mean", "SE", "Lower Bound", "Upper Bound", "$n_{eff}$","$\\hat{R}$")

# The column with row names is added to the table 
fitrstanhiertable3 <- add_column(fitrstanhiertable2, 
                                 Parameters = rownamesfitrstanhiertable, 
                                 .before = "Mean")

# Using the apa_table function from the papaja package the results are presented in an apa table 
apa_table(fitrstanhiertable3,   # Data frame with results 
          row.names = FALSE,    # Default row names are deleted (we set them ourself)
          caption = "Posterior Mean, Standard Error (SE) of the Mean, Lower and 
          Upper Bound of the 95\\% Credible Interval, the Number of Effective 
          Samples, and the $\\hat{R}$ of the General Effect Parameters as 
          estimated by Rstan.", 
          align = c("l", "c", "c", "c", "c", "c", "c"),   # How should results be presented, in the midle of the table or outlined left 
          digits = 3,       # The number of digits behind the ".". 
          placement = "h",  # Place after code 
          escape = FALSE)   # This indicates that the results contain latex code that should be read as latex code. It is important to do this in combination with putting the results to "asis" in the {R} section above.

```

```{r inddevtable1, results = "asis", cache = TRUE, echo = FALSE}
fitrstanhiertable2 <- summary(model_fit,   # model fit 
                              # Select variance parameters (we called them g)
                              pars = c("g", "g2", "g3", "g4", "g5", 
                                       "g6"))$summary[,c("mean", "se_mean", 
                                                         "2.5%", "97.5%", 
                                                         "n_eff", "Rhat")]

# The results are transformed to a dataframe that allows us to plot the results nicely in a table 
fitrstanhiertable22 <- as.data.frame(fitrstanhiertable2)

# We manually set the row names of the table using latex for mathematical symbols 
rownamesfitrstanhiertable2 <-  c("$\\sigma_{\\gamma}^2$","$\\sigma_{\\beta}^2$", "$\\sigma_{\\delta_{7}}^2$", "$\\sigma_{\\delta_{6}}^2$", "$\\sigma_{\\delta_{4}}^2$", "$\\sigma_{\\delta_{3}}^2$")

# The column with row names is added to the table 
colnames(fitrstanhiertable22) <- c("Mean", "SE", "Lower Bound", "Upper Bound", "$n_{eff}$","$\\hat{R}$")

# The column with row names is added to the table 
fitrstanhiertable32 <- add_column(fitrstanhiertable22, 
                                  Parameters = rownamesfitrstanhiertable2, 
                                  .before = "Mean")

# Using the apa_table function from the papaja package the results are presented in an apa table 
apa_table(fitrstanhiertable32,  # Data frame with results
          row.names = FALSE,    # Default row names are deleted (we set them ourself)
          caption = "Posterior Variance, Standard Error (SE) of the Variance, 
          Lower and Upper Bound of the 95\\% Credible Interval, the Number of 
          Effective Samples, and the $\\hat{R}$ of the Variance Parameters           as estimated by Rstan.", 
          align = c("l", "c", "c", "c", "c", "c", "c"),  # How should results be presented, in the midle of the table or outlined left 
          digits = 3,        # The number of digits behind the ".". 
          place = "h",       # Place after code 
          escape = FALSE)    # This indicates that the results contain latex code that should be read as latex code. It is important to do this in combination with putting the results to "asis" in the {R} section above.

```
\newpage

### brms 
```{r loadmodelfits2, show = FALSE, echo = FALSE}
setwd('..')
model_fit <- readRDS(file = "R objects/brms/Truncated/BRMSmodelfit18_03012022.rds")

```

\scriptsize 
```{r geneftable2, results = "asis", cache = TRUE, echo = FALSE}
# Obtain the mean, standard error, 95% credible interval, rhat and number of effective samples for the intercept, and the digit and side parameters 
fitbrmshiertable <- summary(model_fit)[[14]][,c("Estimate", "Est.Error", "l-95% CI", "u-95% CI", "Bulk_ESS", "Rhat")]  # variance

# The results are transformed to a dataframe that allows us to plot the results nicely in a table 
fitbrmshiertable2 <- as.data.frame(fitbrmshiertable)
fitbrmshiertable2 <- rbind(fitbrmshiertable2[6,], fitbrmshiertable2[1:5,])

# We manually set the row names of the table using latex for mathematical symbols 
rownamesfitbrmshiertable <-  c("$\\mu_{\\gamma}$","$\\mu_{\\beta}$", "$\\mu_{\\delta_{7}}$", 
                               "$\\mu_{\\delta_{6}}$", "$\\mu_{\\delta_{4}}$", "$\\mu_{\\delta_{3}}$")

# We manually set the column names of the table 
colnames(fitbrmshiertable2) <- c("Mean", "SE", "Lower Bound", "Upper Bound", "$n_{eff}$","$\\hat{R}$")

# The column with row names is added to the table 
fitbrmshiertable3 <- add_column(fitbrmshiertable2, 
                                 Parameters = rownamesfitbrmshiertable, 
                                 .before = "Mean")

# Using the apa_table function from the papaja package the results are presented in an apa table 
apa_table(fitbrmshiertable3,   # Data frame with results 
          row.names = FALSE,    # Default row names are deleted (we set them ourself)
          caption = "Posterior Mean, Standard Error (SE) of the Mean, Lower and 
          Upper Bound of the 95\\% Credible Interval, the Number of Effective 
          Samples, and the $\\hat{R}$ of the General Effect Parameters as estimated by brms. ", 
          align = c("l", "c", "c", "c", "c", "c", "c"),   # How should results be presented, in the midle of the table or outlined left 
          digits = 3,       # The number of digits behind the ".". 
          placement = "h",  # Place after code 
          escape = FALSE)   # This indicates that the results contain latex code that should be read as latex code. It is important to do this in combination with putting the results to "asis" in the {R} section above.

```
\normalsize

\scriptsize 
```{r inddevtable2, results = "asis", cache = TRUE, echo = FALSE}
# Obtain standard deviation estimates 
fitbrmshiertable2 <- summary(model_fit)[[17]]$sub1[,c("Estimate", "Est.Error", "l-95% CI", "u-95% CI", "Bulk_ESS", "Rhat")]

# The results are transformed to a dataframe that allows us to plot the results nicely in a table 
fitbrmshiertable22 <- as.data.frame(fitbrmshiertable2)
fitbrmshiertable22 <- rbind(fitbrmshiertable22[6,], fitbrmshiertable22[1:5,])

# We manually set the row names of the table using latex for mathematical symbols 
rownamesfitbrmshiertable2 <-  c("$\\sigma_{\\gamma}$","$\\sigma_{\\beta}$", "$\\sigma_{\\delta_{7}}$", 
                                "$\\sigma_{\\delta_{6}}$", "$\\sigma_{\\delta_{4}}$", "$\\sigma_{\\delta_{3}}$")

# The column with row names is added to the table 
colnames(fitbrmshiertable22) <- c("Mean", "SE", "Lower Bound", "Upper Bound", "$n_{eff}$","$\\hat{R}$")

# The column with row names is added to the table 
fitbrmshiertable32 <- add_column(fitbrmshiertable22, 
                                  Parameters = rownamesfitbrmshiertable2, 
                                  .before = "Mean")

# Using the apa_table function from the papaja package the results are presented in an apa table 
apa_table(fitbrmshiertable32,  # Data frame with results
          row.names = FALSE,    # Default row names are deleted (we set them ourself)
          caption = "Posterior Standard Deviation, Standard Error (SE) of the Standard Deviation, 
          Lower and Upper Bound of the 95\\% Credible Interval, the Number of 
          Effective Samples, and the $\\hat{R}$ of the Standard Deviation Parameters           as estimated by brms", 
          align = c("l", "c", "c", "c", "c", "c", "c"),  # How should results be presented, in the midle of the table or outlined left 
          digits = 3,        # The number of digits behind the ".". 
          place = "h",       # Place after code 
          escape = FALSE)    # This indicates that the results contain latex code that should be read as latex code. It is important to do this in combination with putting the results to "asis" in the {R} section above.

```
\normalsize 
\newpage

## Log normal model 
### Rstan 
```{r loadmodelfit3, echo = FALSE}
setwd('..')
logmodel_fit <- readRDS(file = "R objects/rstan/log-normal model/hier_logmodelc_adj13012022.rds")
```

\scriptsize 
```{r geneftable3, results = "asis", cache = TRUE, echo = FALSE}
# Obtain the mean, standard error, 95% credible interval, rhat and number of effective samples for the intercept, and the digit and side parameters 
fitrstanhiertable <- summary(logmodel_fit,  # model fit 
                             # Select parameters 
                             pars = c("mu", "mu2", "mu3", "mu4", "mu5", 
                                      "mu6"))$summary[,c("mean", "se_mean", 
                                                         "2.5%", "97.5%", 
                                                         "n_eff", "Rhat")]

# The results are transformed to a dataframe that allows us to plot the results nicely in a table 
fitrstanhiertable2 <- as.data.frame(fitrstanhiertable)

# We manually set the row names of the table using latex for mathematical symbols 
rownamesfitrstanhiertable <-  c("$\\mu_{\\gamma}$","$\\mu_{\\beta}$", "$\\mu_{\\delta_{7}}$", "$\\mu_{\\delta_{6}}$", "$\\mu_{\\delta_{4}}$", "$\\mu_{\\delta_{3}}$")

# We manually set the column names of the table 
colnames(fitrstanhiertable2) <- c("Mean", "SE", "Lower Bound", "Upper Bound", "$n_{eff}$","$\\hat{R}$")

# The column with row names is added to the table 
fitrstanhiertable3 <- add_column(fitrstanhiertable2, 
                                 Parameters = rownamesfitrstanhiertable, 
                                 .before = "Mean")

# Using the apa_table function from the papaja package the results are presented in an apa table 
apa_table(fitrstanhiertable3,   # Data frame with results 
          row.names = FALSE,    # Default row names are deleted (we set them ourself)
          caption = "Posterior Mean, Standard Error (SE) of the Mean, Lower and 
          Upper Bound of the 95\\% Credible Interval, the Number of Effective 
          Samples, and the $\\hat{R}$ of the General Effect Parameters as 
          estimated by Rstan.", 
          align = c("l", "c", "c", "c", "c", "c", "c"),   # How should results be presented, in the midle of the table or outlined left 
          digits = 3,       # The number of digits behind the ".". 
          placement = "h",  # Place after code 
          escape = FALSE)   # This indicates that the results contain latex code that should be read as latex code. It is important to do this in combination with putting the results to "asis" in the {R} section above.

```
\normalsize

\scriptsize 
```{r inddevtable3, results = "asis", cache = TRUE, echo = FALSE}
fitrstanhiertable2 <- summary(logmodel_fit,   # model fit 
                              # Select variance parameters (we called them g)
                              pars = c("g", "g2", "g3", "g4", "g5", 
                                       "g6"))$summary[,c("mean", "se_mean", 
                                                         "2.5%", "97.5%", 
                                                         "n_eff", "Rhat")]

# The results are transformed to a dataframe that allows us to plot the results nicely in a table 
fitrstanhiertable22 <- as.data.frame(fitrstanhiertable2)

# We manually set the row names of the table using latex for mathematical symbols 
rownamesfitrstanhiertable2 <-  c("$\\sigma_{\\gamma}^2$","$\\sigma_{\\beta}^2$", "$\\sigma_{\\delta_{7}}^2$", "$\\sigma_{\\delta_{6}}^2$", "$\\sigma_{\\delta_{4}}^2$", "$\\sigma_{\\delta_{3}}^2$")

# The column with row names is added to the table 
colnames(fitrstanhiertable22) <- c("Mean", "SE", "Lower Bound", "Upper Bound", "$n_{eff}$","$\\hat{R}$")

# The column with row names is added to the table 
fitrstanhiertable32 <- add_column(fitrstanhiertable22, 
                                  Parameters = rownamesfitrstanhiertable2, 
                                  .before = "Mean")

# Using the apa_table function from the papaja package the results are presented in an apa table 
apa_table(fitrstanhiertable32,  # Data frame with results
          row.names = FALSE,    # Default row names are deleted (we set them ourself)
          caption = "Posterior Variance, Standard Error (SE) of the Variance, 
          Lower and Upper Bound of the 95\\% Credible Interval, the Number of 
          Effective Samples, and the $\\hat{R}$ of the Variance Parameters           as estimated by Rstan.", 
          align = c("l", "c", "c", "c", "c", "c", "c"),  # How should results be presented, in the midle of the table or outlined left 
          digits = 3,        # The number of digits behind the ".". 
          place = "h",       # Place after code 
          escape = FALSE)    # This indicates that the results contain latex code that should be read as latex code. It is important to do this in combination with putting the results to "asis" in the {R} section above.

```
\normalsize 
\newpage

### brms

\scriptsize 
```{r loadmodelfits4, show = FALSE, echo = FALSE}
# Log model brms 
setwd('..')
logmodel_fit <- readRDS(file = "R objects/brms/log-normal/brmmodel8log13012022.rds")

```
\normalsize

\scriptsize 
```{r geneftable4, results = "asis", cache = TRUE, echo = FALSE}
# Obtain the mean, standard error, 95% credible interval, rhat and number of effective samples for the intercept, and the digit and side parameters 
fitbrmshiertable <- summary(logmodel_fit)[[14]][,c("Estimate", "Est.Error", "l-95% CI", "u-95% CI", "Bulk_ESS", "Rhat")]  # variance

# The results are transformed to a dataframe that allows us to plot the results nicely in a table 
fitbrmshiertable2 <- as.data.frame(fitbrmshiertable)

# We manually set the row names of the table using latex for mathematical symbols 
rownamesfitbrmshiertable <-  c("$\\mu_{\\gamma}$","$\\mu_{\\beta}$", "$\\mu_{\\delta_{7}}$", 
                               "$\\mu_{\\delta_{6}}$", "$\\mu_{\\delta_{4}}$", "$\\mu_{\\delta_{3}}$")

# We manually set the column names of the table 
colnames(fitbrmshiertable2) <- c("Mean", "SE", "Lower Bound", "Upper Bound", "$n_{eff}$","$\\hat{R}$")

# The column with row names is added to the table 
fitbrmshiertable3 <- add_column(fitbrmshiertable2, 
                                 Parameters = rownamesfitbrmshiertable, 
                                 .before = "Mean")

# Using the apa_table function from the papaja package the results are presented in an apa table 
apa_table(fitbrmshiertable3,   # Data frame with results 
          row.names = FALSE,    # Default row names are deleted (we set them ourself)
          caption = "Posterior Mean, Standard Error (SE) of the Mean, Lower and Upper Bound of the 95\\% Credible Interval, Number of Effective Samples, and $\\hat{R}$ of the General Effect Parameters as estimated by brms. ", 
          align = c("l", "c", "c", "c", "c", "c", "c"),   # How should results be presented, in the midle of the table or outlined left 
          digits = 3,       # The number of digits behind the ".". 
          placement = "h",  # Place after code 
          escape = FALSE)   # This indicates that the results contain latex code that should be read as latex code. It is important to do this in combination with putting the results to "asis" in the {R} section above.

```
\normalsize

\scriptsize 
```{r inddevtable4, results = "asis", cache = TRUE, echo = FALSE}
# Obtain standard deviation estimates 
fitbrmshiertable2 <- summary(logmodel_fit)[[17]]$sub1[,c("Estimate", "Est.Error", "l-95% CI", "u-95% CI", "Bulk_ESS", "Rhat")]

# The results are transformed to a dataframe that allows us to plot the results nicely in a table 
fitbrmshiertable22 <- as.data.frame(fitbrmshiertable2)

# We manually set the row names of the table using latex for mathematical symbols 
rownamesfitbrmshiertable2 <-  c("$\\sigma_{\\gamma}$","$\\sigma_{\\beta}$", "$\\sigma_{\\delta_{7}}$", 
                                "$\\sigma_{\\delta_{6}}$", "$\\sigma_{\\delta_{4}}$", "$\\sigma_{\\delta_{3}}$")

# The column with row names is added to the table 
colnames(fitbrmshiertable22) <- c("Mean", "SE", "Lower Bound", "Upper Bound", "$n_{eff}$","$\\hat{R}$")

# The column with row names is added to the table 
fitbrmshiertable32 <- add_column(fitbrmshiertable22, 
                                  Parameters = rownamesfitbrmshiertable2, 
                                  .before = "Mean")

# Using the apa_table function from the papaja package the results are presented in an apa table 
apa_table(fitbrmshiertable32,  # Data frame with results
          row.names = FALSE,    # Default row names are deleted (we set them ourself)
          caption = "Posterior Variance, Standard Error (SE) of the Standard Deviation, Lower and Upper Bound of the 95\\% Credible Interval, Number of Effective Samples, and $\\hat{R}$ of the Variance Parameters as estimated by brms.", 
          align = c("l", "c", "c", "c", "c", "c", "c"),  # How should results be presented, in the midle of the table or outlined left 
          digits = 3,        # The number of digits behind the ".". 
          place = "h",       # Place after code 
          escape = FALSE)    # This indicates that the results contain latex code that should be read as latex code. It is important to do this in combination with putting the results to "asis" in the {R} section above.

```
\normalsize